Growth model analysis 3spp WB

### 2019-03-22 MS_Analysis_3spp_4rivers_Patterns_from 2spp.Rmd
added ...bySpp.Rmd

@@

modCSR <- bam(observedLength ~ cohort * species * river + 
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', by = cohortOrdered) +
               s(ydayCumul, bs = 'cr', by = speciesOrdered) +
               s(ydayCumul, bs = 'cr', by = riverOrdered) +
               s(ydayCumul, bs = 'cr', by = cohortRiverOrdered) +
               s(ydayCumul, bs = 'cr', by = speciesCohortOrdered) +
               s(ydayCumul, bs = 'cr', by = speciesRiverOrdered) +
               s(ydayCumul, bs = 'cr', by = cohortSpeciesRiverOrdered), 
             data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
summary(modCSR)
#plot(modCSR)

@@

ggplot(predGrid2CSR, aes(ydayCumul,predCSR, color = species)) +
  geom_point() + 
  geom_line() +
  ylim(50,210) +
  facet_grid(river ~ cohort)

##########################################################
# by cohortSpeciesRiver - additive splines
system.time(
modCSR_add <- bam(observedLength ~ cohort * species * river + 
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', by = cohortOrdered) +
               s(ydayCumul, bs = 'cr', by = speciesOrdered) +
               s(ydayCumul, bs = 'cr', by = riverOrdered) +
               s(ydayCumul, bs = 'cr', by = cohortRiverOrdered) +
               s(ydayCumul, bs = 'cr', by = speciesCohortOrdered) +
               s(ydayCumul, bs = 'cr', by = speciesRiverOrdered), 
             data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
)
summary(modCSR_add)
#plot(modCSR_add)

predGridCSR_add <- expand.grid(cohortSpeciesRiverOrdered = unique(modCSR_add$model$cohortSpeciesRiverOrdered),
                           ydayCumul = seq(250,1000, by = 30)) %>%
                    mutate(cohortOrdered = as.numeric(substr(cohortSpeciesRiverOrdered,1,4)),
                           speciesOrdered = substr(cohortSpeciesRiverOrdered,6,8),
                           riverOrdered = substring(cohortSpeciesRiverOrdered,10),
                           cohortRiverOrdered = paste(cohortOrdered, riverOrdered, sep = '_'),
                           speciesCohortOrdered = paste(speciesOrdered, cohortOrdered, sep = '_'),
                           speciesRiverOrdered = paste(speciesOrdered, riverOrdered, sep = '_'),
                           cohortCentered = cohortOrdered - cohortOffset,
                           cohort = cohortOrdered,
                           species = speciesOrdered,
                           river = riverOrdered)

predCSR_add <- predict(modCSR_add, newdata = predGridCSR_add)#, type = 'lpmatrix')

predGrid2CSR_add <- bind_cols(predGridCSR_add,as.data.frame(predCSR_add))

ggplot(predGrid2CSR_add, aes(ydayCumul,predCSR_add, color = species)) +
  geom_point() + 
  geom_line() +
  ylim(50,210) +
  facet_grid(river ~ cohort)

@@

modSR <- bam(observedLength ~ cohort * species * river + 
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', by = speciesOrdered) +
               s(ydayCumul, bs = 'cr', by = riverOrdered) +
               s(ydayCumul, bs = 'cr', by = speciesRiverOrdered), 
             data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
summary(modSR)
#plot(modSR, scale = 0)

@@

ggplot(predGrid2SR, aes(ydayCumul,predSR, color = species)) +
  geom_point() + 
  geom_line() +
  facet_grid(river ~ cohortOrdered)

# by speciesCohort
system.time(
modSC <- bam(observedLength ~ cohort * species * river + 
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', by = cohortOrdered) +
               s(ydayCumul, bs = 'cr', by = speciesOrdered) +
               s(ydayCumul, bs = 'cr', by = speciesCohortOrdered), 
             data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
)
summary(modSC)
#plot(modSC, scale = 0)

@@

ggplot(predGrid2SC, aes(ydayCumul,predSC, color = species)) +
  geom_point() + 
  geom_line() +
  facet_grid(river ~ cohortOrdered)

# by cohortRiver

modCR <- bam(observedLength ~ cohort * species * river + 
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', by = cohortOrdered) +
               s(ydayCumul, bs = 'cr', by = riverOrdered) +
               s(ydayCumul, bs = 'cr', by = cohortRiverOrdered), 
               data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
summary(modCR)
#plot(modCR, scale = 0)

@@

modCSRY <- bam(observedLength ~ cohort * species * river + 
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', by = cohortOrdered) +
               s(ydayCumul, bs = 'cr', by = speciesOrdered) +
               s(ydayCumul, bs = 'cr', by = riverOrdered) +
               s(ydayCumul, bs = 'cr', by = cohortRiverOrdered) +
               s(ydayCumul, bs = 'cr', by = speciesCohortOrdered) +
               s(ydayCumul, bs = 'cr', by = speciesRiverOrdered) +
               s(ydayCumul, bs = 'cr', by = cohortSpeciesRiverOrdered) +
               #s(year, bs = 'cr', by = speciesOrdered), 
               s(year, bs = 'cr', by = speciesFactor),
               data = vB %>% filter(cohort %in% 2000:2013, species %in% c('bkt','bnt')), 
               method = "REML")
summary(modCSRY)
#plot(modCSRY, select = 189)

@@

ggplot(predGrid2CSRY, aes(ydayCumul, predCSRY, color = species)) +
  geom_point() + 
  geom_line() +
  ylim(50,300) +
  facet_grid(river ~ cohortOrdered)

@@

mod <- bam(observedLength ~ cohort * species * river + 
               s(ydayCumul, bs = 'cr'), 
               data = vB %>% filter(cohort %in% 2000:2013, species %in% c('bkt','bnt')), 
               method = "REML")
summary(modCSRY)

save(mod,modCR,modCSR,modSC,modSR,modCSRY, file = './data/out/mods.RData')
AIC(modCSR,modSR,modSC,modCR,mod)  %>% rownames_to_column() %>% arrange(AIC)

>>>>

kIn <- 4
dataIn <- vB %>% filter(cohort %in% 2000:2013, species %in% c('bkt','bnt'))
```

``` {r gam models}

@@

modCSR <- bam(observedLength ~ cohort * species * river + 
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = speciesOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = speciesCohortOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = speciesRiverOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortSpeciesRiverOrdered), 
             data = dataIn, method = "REML")
summary(modCSR)
#plot(modCSR)

@@

ggplot(predGrid2CSR, aes(ydayCumul,predCSR, color = species)) +
  geom_point() + 
  geom_line() +
  ylim(50,210) +
  facet_grid(river ~ cohort)+
  ggtitle('modCSR')

##########################################################
# by cohortSpeciesRiver - additive splines, no 3-way spline
system.time(
modC_S_R <- bam(observedLength ~ cohort * species * river + 
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = speciesOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = speciesCohortOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = speciesRiverOrdered), 
             data = dataIn, method = "REML")
)
summary(modC_S_R)
#plot(modC_S_R)

predGridC_S_R <- expand.grid(cohortSpeciesRiverOrdered = unique(modCSR$model$cohortSpeciesRiverOrdered), # use modCSR to get 3-way
                               ydayCumul = seq(250,1000, by = 30)) %>%
                    mutate(cohortOrdered = as.numeric(substr(cohortSpeciesRiverOrdered,1,4)),
                           speciesOrdered = substr(cohortSpeciesRiverOrdered,6,8),
                           riverOrdered = substring(cohortSpeciesRiverOrdered,10),
                           cohortRiverOrdered = paste(cohortOrdered, riverOrdered, sep = '_'),
                           speciesCohortOrdered = paste(speciesOrdered, cohortOrdered, sep = '_'),
                           speciesRiverOrdered = paste(speciesOrdered, riverOrdered, sep = '_'),
                           cohortCentered = cohortOrdered - cohortOffset,
                           cohort = cohortOrdered,
                           species = speciesOrdered,
                           river = riverOrdered)

predC_S_R <- predict(modC_S_R, newdata = predGridC_S_R)#, type = 'lpmatrix')

predGrid2C_S_R <- bind_cols(predGridC_S_R,as.data.frame(predC_S_R))

ggplot(predGrid2C_S_R, aes(ydayCumul,predC_S_R, color = species)) +
  geom_point() + 
  geom_line() +
  ylim(50,210) +
  facet_grid(river ~ cohort) +
  ggtitle('modC_S_R')

@@

modSR <- bam(observedLength ~ cohort * species * river + 
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = speciesOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = speciesRiverOrdered), 
             data = dataIn, method = "REML")
summary(modSR)
#plot(modSR, scale = 0)

@@

ggplot(predGrid2SR, aes(ydayCumul,predSR, color = species)) +
  geom_point() + 
  geom_line() +
  facet_grid(river ~ cohortOrdered)+
  ggtitle('modSR')

#######################################
# by speciesCohort
system.time(
modSC <- bam(observedLength ~ cohort * species * river + 
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = speciesOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = speciesCohortOrdered), 
             data = dataIn, method = "REML")
)
summary(modSC)
#plot(modSC, scale = 0)

@@

ggplot(predGrid2SC, aes(ydayCumul,predSC, color = species)) +
  geom_point() + 
  geom_line() +
  facet_grid(river ~ cohortOrdered) +
  ggtitle('modSC')

##################################
# by cohortRiver

modCR <- bam(observedLength ~ cohort * species * river + 
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered), 
               data = dataIn, method = "REML")
summary(modCR)
#plot(modCR, scale = 0)

@@

modCSRY <- bam(observedLength ~ cohort * species * river + 
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = speciesOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = speciesCohortOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = speciesRiverOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortSpeciesRiverOrdered) +
               #s(year, bs = 'cr', by = speciesOrdered), 
               s(year, bs = 'cr', k = kIn, by = speciesFactor),
               data = dataIn, 
               method = "REML")
summary(modCSRY)
plot(modCSRY, pages=10, scale = 0)

@@

ggplot(predGrid2CSRY, aes(ydayCumul, predCSRY, color = species)) +
  geom_point() + 
  geom_line() +
  ylim(50,300) +
  facet_grid(river ~ cohortOrdered) +
  ggtitle('modCSRY')

# trying a couple of models
modCSRC <- bam(observedLength ~ cohort * species * river + 
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = speciesOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = speciesCohortOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = speciesRiverOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortSpeciesRiverOrdered) +
               #s(year, bs = 'cr', by = speciesOrdered), 
               s(cohort, bs = 'cr', k = kIn, by = speciesFactor),
               data = dataIn, 
               method = "REML")
summary(modCSRC)
plot(modCSRC, pages=8, scale = 0)

modC_S_RC <- bam(observedLength ~ cohort + species + river + 
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = speciesOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = speciesCohortOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = speciesRiverOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortSpeciesRiverOrdered) +
               s(cohort, bs = 'cr', k = kIn, by = speciesFactor),
               data = dataIn, 
               method = "REML")
summary(modC_S_RC)
plot(modCSRC, pages=8, scale = 0)

# modCSR_Ind <- bam(observedLength ~ cohort * species * river + 
#                s(ydayCumul, bs = 'cr') +
#                s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
#                s(ydayCumul, bs = 'cr', k = kIn, by = speciesOrdered) +
#                s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
#                s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
#                s(ydayCumul, bs = 'cr', k = kIn, by = speciesCohortOrdered) +
#                s(ydayCumul, bs = 'cr', k = kIn, by = speciesRiverOrdered) +
#                s(ydayCumul, bs = 'cr', k = kIn, by = cohortSpeciesRiverOrdered) +
#                s(ydayCumul, tagIndexFactor, bs="fs", m=1, k = kIn),
#                data = dataIn, 
#                method = "REML")
# summary(modCSR_Ind)

@@

mod <- bam(observedLength ~ cohort * species * river + 
               s(ydayCumul, bs = 'cr'), 
               data = dataIn, 
               method = "REML")
summary(mod)

predGrid <- expand.grid(cohortSpeciesRiverOrdered = unique(modCSR$model$cohortSpeciesRiverOrdered),
                        ydayCumul = seq(250,1000, by = 30)) %>%
                    mutate(cohortOrdered = as.numeric(substr(cohortSpeciesRiverOrdered,1,4)),
                           speciesOrdered = substr(cohortSpeciesRiverOrdered,6,8),
                           riverOrdered = substring(cohortSpeciesRiverOrdered,10),
                           cohort = cohortOrdered,
                           species = speciesOrdered,
                           river = riverOrdered)

pred <- predict(mod, newdata = predGrid)#, type = 'lpmatrix')

ggplot(predGrid2, aes(ydayCumul, pred, color = species)) +
  geom_point() + 
  geom_line() +
  facet_grid(river ~ cohortOrdered) +
  ggtitle('mod')

##########################################

save(mod,modCR,modCSR,modSC,modSR,modCSRY,modC_S_R, file = './data/out/mods.RData')
AIC(modC_S_RC,modC_S_R,modCSR,modSR,modSC,modCR,mod)  %>% rownames_to_column() %>% arrange(AIC)
